Numerical modelling of the Cordilleran ice sheet

### 2014-07-19 figures/model-pdd.tex
Rework PDD model sketch.

\documentclass[border=5mm,tikz]{standalone}
\usepackage{amsmath}
\usetikzlibrary{chains}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes}

\newcommand{\PDD}{\ensuremath{\text{PDD}}}
\newcommand{\Hs}{\ensuremath{H_{\text{snow}}}}
\newcommand{\Ms}{\ensuremath{\text{Melt}_{\text{snow}}}}
\newcommand{\Mi}{\ensuremath{\text{Melt}_{\text{ice}}}}

\begin{document}
  \begin{tikzpicture}[ auto, thick, font=\small, node distance=3mm,
    block/.style={rectangle, rounded corners, draw, fill=gray!5, text width=2in, text centered,
      minimum width=2in, minimum height=1cm},
    big block/.style={block, text width=3in},
    decision/.style={diamond, draw, fill=gray!5},
    line/.style={draw, -latex'},
    start chain=1 going below, every node/.style={on chain},
    every join/.style={-latex'},
    ]

    \node [big block] {Compute snow accumulation \mbox{``Accumulation''}, discarding rain};
    \node [big block, join] {Compute the number of \mbox{positive} degree days (\PDD) };
    \node [big block, join] {Update snow depth \mbox{$\Hs \gets \Hs + \text{Accumulation}$}};
    \node [big block, join] {Compute maximum possible snow melt \mbox{$\Ms^{\text{max}} := \PDD\times F_{s}$}};
    \node (melt all snow) [decision, join] {$\Ms^{\text{max}} > \Hs?$};

    \begin{scope}[start branch=yes, node distance=4em]
      \node (ice melt case) [block, on chain=going right] {\mbox{Melt all available snow:} $\Ms := \Hs$};
    \end{scope}
    \begin{scope}[start branch=no, node distance=4em]
      \node (no ice melt case) [block, on chain=going left] {Melt some of the snow: \mbox{$\Ms := \Ms^{\text{max}}$}};
    \end{scope}

    \node (excess pdds) [big block] {$\PDD_{\text{excess}} := \PDD - \Ms / F_{s}$};
    \node [big block, join] {Melt some ice: \mbox{$\Mi := \PDD_{\text{excess}} \times F_{i}$}};
    \node (refreeze ice) [decision, join] {Refreeze melted ice?};

    \begin{scope}[start branch=yes, node distance=4em]
      \node (refreeze both) [block, on chain=going right] {Refreeze is \mbox{$(\Ms + \Mi) \times \theta_{\text{refreeze}}$}};
    \end{scope}

    \begin{scope}[start branch=no, node distance=4em]
      \node (refreeze snow) [block, on chain=going left] {Refreeze is $\Ms \times \theta_{\text{refreeze}}$};
    \end{scope}

    \node (runoff) [big block] {$\text{Runoff} := \Ms + \Mi - \text{Refreeze}$};
    \node [big block, join] {Compute surface mass balance: $\text{SMB} := \text{Accumulation} - \text{Runoff}$};
    \node [big block, join] {Update snow depth: $\Hs \gets \Hs - \Ms$};
    % special edges:
    \path[line] (melt all snow) -- node {Yes} (ice melt case);
    \path[line] (melt all snow) -- node {No}  (no ice melt case);
    \path[line] (ice melt case) |- (excess pdds);
    \path[line] (no ice melt case) |- (excess pdds);
    \path[line] (refreeze ice) -- node {Yes} (refreeze both);
    \path[line] (refreeze ice) -- node {No}  (refreeze snow);
    \path[line] (refreeze both) |- (runoff);
    \path[line] (refreeze snow) |- (runoff);
  \end{tikzpicture}

\end{document}

>>>>

\documentclass[border=0mm,tikz]{standalone}
\usetikzlibrary{shapes}
\usetikzlibrary{calc}

% TikZ styles
\tikzstyle{block} = [draw, rectangle, fill=gray!5, rounded corners,
                     text centered, minimum height=1cm]
\tikzstyle{trunk}  = [block, text width=55mm, minimum width=60mm]
\tikzstyle{branch} = [block, text width=32.5mm, minimum width=35mm]
\tikzstyle{split} = [draw, diamond, fill=gray!5, minimum size=2.5cm]
\newcommand{\PDD}{\mathrm{PDD}}
\newcommand{\SMB}{\mathrm{SMB}}

% begin document
\begin{document}
\begin{tikzpicture}[thick, font=\small, >=latex]

% background
\fill[white] (-0.25,-0.25) rectangle +(8.5,14.75);
%\draw[help lines, lightgray] (0,0) grid (8,14);

% add nodes
\path     (4, 13.75) node [trunk] (acc) {Compute snow accumulation, $A_s$}
  -- ++(0, -1.5) node [trunk] (pdd) {Compute PDDs}
  -- ++(0, -1.5) node [trunk]  (s1) {Update snow depth \\
                                      $S \gets S + A_s$}
  -- ++(0, -1.5) node [trunk] (msp) {Compute potential snow melt \\
                                      $M_{s,p} := \PDD \times F_{s}$}
  -- ++(0, -2.25) node [split] (split) {$M_{s,p} > H_s?$};

\path ($(split)+(-2,-2)$) node [branch] (no)  {Melt some snow \\
                                               $M_s := M_{s,p}$}
  -- ++(0, -1.5) node [branch] (nomi) {Melt no ice \\
                                      $M_i := 0$};

\path ($(split)+(+2,-2)$) node [branch] (yes) {Melt all snow \\
                                               $M_s := H_s$}
  -- ++(0, -1.5) node [branch] (yesmi) {Melt some ice \\
                                        $M_i := (\PDD - \frac{M_s}{F_s}) \times F_i$};

\path ($(split)+(0,-5)$) node [trunk](smb) {Compute surf. mass balance \\
                                      $\SMB := A_s - M_s - M_i$}
  -- ++(0, -1.5) node [trunk]  (s2) {Update snow depth \\
                                      $S \gets S - M_s$};

% add connections
\draw[->] (acc) -- (pdd);
\draw[->] (pdd) -- (s1);
\draw[->] (s1) -- (msp);
\draw[->] (msp) -- (split);
\draw[->] (split) -| (no) node [pos=0.3, above] {No};
\draw[->] (no) -- (nomi);
\draw[->] (nomi) -- (smb);
\draw[->] (split) -| (yes) node [pos=0.3, above] {Yes};
\draw[->] (yes) -- (yesmi);
\draw[->] (yesmi) -- (smb);
\draw[->] (smb) -- (s2);

\end{tikzpicture}
\end{document}

### 2014-07-19 kappa.tex
Rework PDD model sketch.

At the interface between the ice sheet and the atmosphere
(\cref{fig:model-interfaces}),
surface mass fluxes are computed from monthly mean surface air temperature,
$T_m$, and monthly precipitation, $P_m$, using a temperature-index model
\citep[e.g.,][]{Hock.2003}. Surface accumulation, $a_s$, equals precipitation
when temperature is below ${T_0=0\,\degC}$, and decreases to zero\renote{
    Well spotted, thanks!}
linearly with temperature between $T_0$ and ${T_1=2\,\degC}$,
\begin{equation}
    a_s =
    \begin{cases}
        P_m     & \text{if}\ T_m \le T_0 \,, \\
        P_m \frac{T_1-T_m}{T_1-T_0}
                & \text{if}\ T_0 < T_m < T_1 \,, \\
        0       & \text{if}\ T_1 \le T_m \,,
    \end{cases}
\end{equation}

Surface melt, $m_s$, is assumed proportional to the number of positive degree
days (PDD), defined over an arbitrary time interval, $[t_1, t_2]$, as the
integral of positive Celcius temperature $T-T_0$,
\begin{equation}
    \mathrm{PDD} = \int_{0}^{A}\max(T(t)-T_0,0)\,dt \,.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{model-pdd}
  \caption{PDD model. From \citet{PISM-authors.2014}
           \todo{This is again directly from the PISM manual. I can simplify
                 it by removing the unused refreezing branch, I will match
                 notations with the text and refer to equations in use.}}
  \label{fig:model-pdd}
\end{figure}

>>>>

At the interface between the ice sheet and the atmosphere
(\cref{fig:model-interfaces}),
surface mass fluxes are computed from monthly mean surface air temperature,
$T_m$, and monthly precipitation, $P_m$, using a temperature-index model
\citep[\cref{fig:model-pdd}; cf.][]{Hock.2003}. Snow accumulation, $A_s$,
equals precipitation
when temperature is below ${T_0=0\,\degC}$, and decreases to zero\renote{
    Well spotted, thanks!}
linearly with temperature between $T_0$ and ${T_1=2\,\degC}$,
\begin{equation}
    A_s =
    \begin{cases}
        P_m     & \text{if}\ T_m \le T_0 \,, \\
        P_m \frac{T_1-T_m}{T_1-T_0}
                & \text{if}\ T_0 < T_m < T_1 \,, \\
        0       & \text{if}\ T_1 \le T_m \,,
    \end{cases}
\end{equation}

Surface melt, $M$, is assumed proportional to the number of positive degree
days (PDD), defined over an arbitrary time interval, $[t_1, t_2]$, as the
integral of positive Celcius temperature $T-T_0$,
\begin{equation}
    \mathrm{PDD} = \int_{0}^{A}\max(T(t)-T_0,0)\,dt \,.
\end{equation}

@@

\begin{figure}
  \centering
  \includegraphics{model-pdd}
  \caption{Positive degree-day algorithm used to compute accumulation and melt
           at the ice sheet surface. Adapted from \citet{PISM-authors.2014}.}
  \label{fig:model-pdd}
\end{figure}
